Enforced Invariant Confluence

### 2018-04-22 doc/paper/sections/interactive_decision_procedure.tex
Finished merge reduction section.
Documented segmented invariant-confluence DP.
Shrunk eval.
Formatting tweaks.

\section{An Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsf{Helper}}
\newcommand{\IsIclosed}{\textsf{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{Interactive Invariant-Confluence Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
    \State
    \Return $I(s_0)$ and
    \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

    \algocomment{$R$ is a set of $\sTIreachable{}$ states.}
    \algocomment{$NR$ is a set of $\sTIunreachable{}$ states.}
    \algocomment{$I(s_0)$ is a precondition.}
    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
    \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
    \If {closed}
      \State \Return true
    \EndIf
    \State Augment $R$ and $NR$ with random search and user input
    \If{$s_1, s_2 \in R$}
    \Return false
    \Else\
    \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
    \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\IsInvConfluent{} first checks that $s_0$ satisfies the invariant. $s_0$ is
reachable, so if it does not satisfy the invariant, then $O$ is not
\sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
and a set $NR$ of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
iteratively increased which removes unreachable states from $I$ until $I
\subseteq \setst{s}{\sTIreachable{s}}$.

First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}.
%
If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
\thmref{IclosureImpliesIconfluence}, $O$ is \sticonfluent{s_0}{T}{I-NR}, so
\[
  \setst{s \in S}{\stireachable{s_0}{T}{I-NR}{s}}
    \subseteq I - NR
    \subseteq I
\]
If $NR$ only contains $\sTIunreachable{}$ states, then the set of
$\sTIreachable{}$ states is equal to set of $\stireachable{s_0}{T}{I-NR}{}$
states which, as we just showed, is a subset of $I$.  Thus, $O$ is
\sTIconfluent{}, so \Helper{} return true.

@@

After $s_1$ and $s_2$ have been labelled as $\sTIreachable{}$ or not, we
continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1 \join
s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{}, so
\Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
$\sTIunreachable{}$, so we recurse.

@@

\textbf{A user can label an unreachable state as reachable.}
In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and return
false, concluding that $O$ is not \sTIconfluent{} even when it is.  This is
inconvenient, but not catastrophic. We can modify
\algoref{InteractiveDecisionProcedure} so that \Helper{} requires that whenever
a user labels a state $s$ as $\sTIreachable{}$, they must also provide an
$\sTIreachable{}$ expression $e$ that evaluates to $s$. Here, $e$ acts a proof
that certifies that $s$ is indeed reachable. This increases the burden on the
user but completely eliminates this class of errors.

\textbf{A user can label a reachable state as unreachable.}
In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
\sTIconfluent. Thus, a user might falsely believe their distributed object to
be \sTIconfluent{} even though it isn't, and eventually one replica of their
distributed object might enter a state that violates the invariant.
%
We can mitigate this in two ways. First, we can aggressively expand $R$. If a
user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
it might not be \sTIconfluent{}). Thus, when a user replicates their
distributed object across a set of servers, they can deploy with the invariant
$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
$NR$ is incorrect and contains some $\sTIreachable{}$ states, then some of
these states are artificially made unreachable, but the system is still
guaranteed to never produce a state which violates $I$.

>>>>

\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsf{Helper}}
\newcommand{\IsIclosed}{\textsf{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{Interactive invariant-confluence decision procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \algocomment{Return if $O$ is \sTIconfluent{}.}
    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
      \State
        \Return $I(s_0)$ and
        \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

    \algocomment{$R$ is a set of \sTIreachable{} states.}
    \algocomment{$NR$ is a set of \sTIunreachable{} states.}
    \algocomment{$I(s_0)$ is a precondition.}
    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
      \If {closed}
        \State \Return true
      \EndIf
      \State Augment $R$ and $NR$ with random search and user input
      \If{$s_1, s_2 \in R$}
        \Return false
      \Else\
        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\IsInvConfluent{} first checks that $s_0$ satisfies the invariant. $s_0$ is
reachable, so if it does not satisfy the invariant, then $O$ is not
\sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
iteratively increased which removes unreachable states from $I$ until $I
\subseteq \setst{s}{\sTIreachablepredicate{s}}$.

First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}.
%
If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
\thmref{IclosureImpliesIconfluence}, $O$ is \sticonfluent{s_0}{T}{I-NR}, so
\[
  \setst{s \in S}{\stireachablepredicate{s_0}{T}{I-NR}{s}}
    \subseteq I - NR
    \subseteq I
\]
If $NR$ only contains \sTIunreachable{} states, then the set of
\sTIreachable{} states is equal to set of \stireachable{s_0}{T}{I-NR}{}
states which, as we just showed, is a subset of $I$.  Thus, $O$ is
\sTIconfluent{}, so \Helper{} return true.

@@

After $s_1$ and $s_2$ have been labelled as \sTIreachable{} or not, we
continue. If both $s_1$ and $s_2$ are \sTIreachable{}, then so is $s_1 \join
s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{}, so
\Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
\sTIunreachable{}, so we recurse.

@@

\textbf{A user can label an unreachable state as reachable.}
In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and return
false, concluding that $O$ is not \sTIconfluent{} even when it is.  This is
inconvenient, but not catastrophic. We can modify
\algoref{InteractiveDecisionProcedure} so that \Helper{} requires that whenever
a user labels a state $s$ as \sTIreachable{}, they must also provide an
\sTIreachable{} expression $e$ that evaluates to $s$. Here, $e$ acts a proof
that certifies that $s$ is indeed reachable. This increases the burden on the
user but completely eliminates this class of errors.

\textbf{A user can label a reachable state as unreachable.}
In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
\sTIconfluent. Thus, a user might falsely believe their distributed object to
be \sTIconfluent{} even though it isn't, and eventually one replica of their
distributed object might enter a state that violates the invariant.
%
We can mitigate this in two ways. First, we can aggressively expand $R$. If a
user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
it might not be \sTIconfluent{}). Thus, when a user replicates their
distributed object across a set of servers, they can deploy with the invariant
$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
$NR$ is incorrect and contains some \sTIreachable{} states, then some of
these states are artificially made unreachable, but the system is still
guaranteed to never produce a state which violates $I$.

### 2018-04-22 doc/paper/sections/introduction.tex
Formatting tweaks.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of invariant-confluence as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
invariant-confluent with respect to an invariant, we can replicate it with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency. However, the task of identifying whether or not an
object actually is invariant-confluent remains a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
invariant-confluent or not, but ideally we would have a general purpose program
that can automatically determine invariant-confluence for us. \textbf{The first
main thrust of this paper is to make progress towards a general-purpose
invariant-confluence decision procedure.}

Designing a general-purpose invariant-confluence decision procedure is
difficult\footnote{Actually, it's impossible. Determining whether an object is
invariant-confluent is, in general, undecidable. Still, we can develop a
decision procedure that works well in the common case.}. Many existing
approaches instead develop a decision procedure for invariant-closure, a
sufficient (but not necessary) condition for
invariant-confluence~\cite{li2012making, li2014automating}. The existing
approaches check to see if an object is invariant-closed. If it is, then they
conclude that it is also invariant-confluent. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
invariant-confluent. In this paper, we take a step back and study the
underlying reason \emph{why} invariant-closure is not necessary for
invariant-confluence. We find that invariant-confluence is fundamentally a
property about reachability, about whether certain values of our replicated
object can actually be obtained in the execution of a system. We discover that
unreachable, yet invariant-satisfying, states cause invariant-closure to be an
unnecessary condition. Using this understanding, we construct a set of modest
conditions under which invariant-closure and invariant-confluence are in fact
\emph{equivalent}, allowing us to reduce the problem of determining
invariant-confluence to that of determining invariant-closure.

Then, we use these conditions to design a general-purpose interactive
invariant-confluence decision procedure. Users iteratively interact with the
decision procedure, eliminating unreachable states from the invariant, moving
towards the conditions under which invariant-closure and invariant-confluence
are equivalent. Meanwhile, the decision-procedure generates example reachable
and unreachable states which help the user recognize patterns that describe
reachability. We also leverage our intuitions about reachability to develop a
new sufficient condition for invariant-confluence, dubbed merge-reducibility.
Merge-reducibility covers some cases that are not covered by invariant-closure,
and it can be automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to take a step beyond
invariant-confluence and develop a generalization of invariant-confluence
called segmented invariant-confluence.} While invariant-confluence
characterizes objects which can be replicated \emph{without any} coordination,
segmented invariant-confluence allows us to replicate non-invariant-confluent
objects with \emph{minimal coordination}. The main idea is that
invariant-satisfying states are divided into a number of segments, and the set
of allowable of transactions within each segment is restricted. Within a
segment, servers act without any coordination, synchronizing only to transition
across segment boundaries. This design highlights the trade-off between
application complexity and coordination-freedom; more complex applications
require more segments which require more coordination, and vice-versa.

In closing, here is an outline of the paper and of our contributions:
\begin{itemize}
  \item
    We propose a novel expression-oriented definition of invariant-confluence
    that is both formal and simple (\secref{InvariantConfluence}).

@@

  \item
    We exploit these conditions to design a general-purpose interactive
    decision procedure for invariant-confluence
    (\secref{InteractiveDecisionProcedure}).

  \item
    We again exploit these conditions to design a novel non-trivial sufficient
    condition for invariant-confluence, merge-reducibility, that does not
    depend on user interaction (\secref{MergeReduction}).

  \item
    We present segmented invariant-confluence: a generalization of
    invariant-confluence that uses a small amount of coordination to maintain
    invariants for replicated objects that are otherwise not
    invariant-confluent (\secref{SegmentedInvariantConfluence}).

  \item
    We implement our decision procedures and sufficient conditions for both
    invariant-confluence and segmented invariant-confluence in a system called
    Lucy, then use Lucy to evaluate the efficacy or our methods
    (\secref{Evaluation}).
\end{itemize}

>>>>

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of \invariantconfluence{} as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
invariant-confluent with respect to an invariant, we can replicate it with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency. However, the task of identifying whether or not an
object actually is invariant-confluent remains a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
invariant-confluent or not, but ideally we would have a general purpose program
that can automatically determine \invariantconfluence{} for us. \textbf{The first
main thrust of this paper is to make progress towards a general-purpose
\invariantconfluence{} decision procedure.}

Designing a general-purpose \invariantconfluence{} decision procedure is
difficult\footnote{Actually, it's impossible. Determining whether an object is
invariant-confluent is, in general, undecidable. Still, we can develop a
decision procedure that works well in the common case.}. Many existing
approaches instead develop a decision procedure for invariant-closure, a
sufficient (but not necessary) condition for
\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
approaches check to see if an object is invariant-closed. If it is, then they
conclude that it is also invariant-confluent. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
invariant-confluent. In this paper, we take a step back and study the
underlying reason \emph{why} invariant-closure is not necessary for
\invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
property about reachability, about whether certain values of our replicated
object can actually be obtained in the execution of a system. We discover that
unreachable, yet invariant-satisfying, states cause invariant-closure to be an
unnecessary condition. Using this understanding, we construct a set of modest
conditions under which invariant-closure and \invariantconfluence{} are in fact
\emph{equivalent}, allowing us to reduce the problem of determining
\invariantconfluence{} to that of determining invariant-closure.

Then, we use these conditions to design a general-purpose interactive
\invariantconfluence{} decision procedure. Users iteratively interact with the
decision procedure, eliminating unreachable states from the invariant, moving
towards the conditions under which invariant-closure and \invariantconfluence{}
are equivalent. Meanwhile, the decision-procedure generates example reachable
and unreachable states which help the user recognize patterns that describe
reachability. We also leverage our intuitions about reachability to develop a
new sufficient condition for \invariantconfluence{}, dubbed merge-reducibility.
Merge-reducibility covers some cases that are not covered by invariant-closure,
and it can be automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to take a step beyond
  \invariantconfluence{} and develop a generalization of \invariantconfluence{}
called segmented \invariantconfluence{}.} While \invariantconfluence{}
characterizes objects which can be replicated \emph{without any} coordination,
segmented \invariantconfluence{} allows us to replicate non-invariant-confluent
objects with \emph{minimal coordination}. The main idea is that
invariant-satisfying states are divided into a number of segments, and the set
of allowable of transactions within each segment is restricted. Within a
segment, servers act without any coordination, synchronizing only to transition
across segment boundaries. This design highlights the trade-off between
application complexity and coordination-freedom; more complex applications
require more segments which require more coordination, and vice-versa.

In closing, here is an outline of the paper and of our contributions:
\begin{itemize}
  \item
    We propose a novel expression-oriented definition of \invariantconfluence{}
    that is both formal and simple (\secref{InvariantConfluence}).

@@

  \item
    We exploit these conditions to design a general-purpose interactive
    decision procedure for \invariantconfluence{}
    (\secref{InteractiveDecisionProcedure}).

  \item
    We again exploit these conditions to design a novel non-trivial sufficient
    condition for \invariantconfluence{}, merge-reducibility, that does not
    depend on user interaction (\secref{MergeReduction}).

  \item
    We present segmented \invariantconfluence{}: a generalization of
    \invariantconfluence{} that uses a small amount of coordination to maintain
    invariants for replicated objects that are otherwise not
    invariant-confluent (\secref{SegmentedInvariantConfluence}).

  \item
    We implement our decision procedures and sufficient conditions for both
    \invariantconfluence{} and segmented \invariantconfluence{} in a system called
    Lucy, then use Lucy to evaluate the efficacy or our methods
    (\secref{Evaluation}).
\end{itemize}

### 2018-04-22 doc/paper/sections/invariant_closure.tex
Formatting tweaks.

\section{Invariant-Closure}\seclabel{InvariantClosure}
Our ultimate goal is to write a program that can automatically decide whether a
given distributed object $O$ is \sTIconfluent{}. Such a program has to
automatically prove or disprove that every reachable state satisfies the
invariant. However, automatically reasoning about the possibly infinite set of
reachable states is challenging, especially because transactions and merge
functions can be complex and can be interleaved arbitrarily in an execution.
Due to this complexity, existing systems that aim to automatically decide
invariant-confluence instead focus on deciding a sufficient condition for
invariant-confluence---dubbed \defword{invariant-closure}---that is simpler to
reason about~\cite{li2012making, li2014automating}. In this section, we define
invariant-closure and study why the condition is sufficient but not necessary.
Armed with this understanding, we present conditions under which it is both
sufficient and necessary.

We say an object $O = (S, \join)$ is \defword{invariant-closed} with respect to
an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
$I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.

@@

\thmref{IclosureImpliesIconfluence} states that invariant-closure is
sufficient for invariant-confluence. Intuitively, our system model ensures that
transaction execution preserves the invariant, so if merging states also
preserves the invariant and if our start state satisfies the invariant, then
inductively it is impossible for us to reach a state that doesn't satisfy the
invariant.

This is good news because checking if an object is invariant-closed is more
straightforward than checking if it is invariant-confluent. Existing systems
typically use an SMT solver like Z3 to check if an object is
invariant-closed~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
is, then by \thmref{IclosureImpliesIconfluence}, it is invariant-confluent.
Unfortunately, invariant-closure is \emph{not} necessary for
invariant-confluence, so if an object is \emph{not} invariant-closed, these
systems cannot conclude that the object is \emph{not} invariant-confluent. The
reason why invariant-closure is not necessary for invariant-confluence is best
explained through an example.

@@

  The invariant and the set of reachable states are illustrated in \figref{Z2}
  in which we draw each state $(x, y)$ as a point in space. The invariant
  consists of the second and fourth quadrant, and the reachable states consist
  of the fourth quadrant. From this, it's immediate that the reachable states
  are a subset of the invariant, so $O$ is invariant-confluent. However,
  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
  invariant-closed. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
  (1, 1)$, we see $\lnot I(s_3)$.
\end{example}

@@

In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
invariant-closed, but $s_1$ is not reachable. This is not particular to
\exampleref{Z2}. In fact, it is fundamentally the reason why invariant-closure
is not equivalent to invariant-confluence. Invariant-confluence is, at its
core, a property of reachable states, but invariant-closure is completely
ignorant of reachability. As a result, invariant-satisfying yet unreachable
states like $s_1$ are the key hurdle preventing invariant-closure from being
equivalent to invariant-confluence. This is formalized by
\thmref{IclosureEquivalentIconfluence}

\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$. If the invariant is a subset of the
  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachable{s}}$), then
  \[
    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
  \]
\end{theorem}
\begin{proof}
  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
  \join s_2$, satisfy the invariant.
\end{proof}

% - define invariant-closure and introduce it as a sufficient condition
% - cite existing work to show that people have used invariant closure to
%   reason about invariant-confluence
% - explain why invariant closure can fail and why its not sufficient using the
%   two-integers example
% - state main theorem

>>>>

\section{Invariant Closure}\seclabel{InvariantClosure}
Our ultimate goal is to write a program that can automatically decide whether a
given distributed object $O$ is \sTIconfluent{}. Such a program has to
automatically prove or disprove that every reachable state satisfies the
invariant. However, automatically reasoning about the possibly infinite set of
reachable states is challenging, especially because transactions and merge
functions can be complex and can be interleaved arbitrarily in an execution.
Due to this complexity, existing systems that aim to automatically decide
\invariantconfluence{} instead focus on deciding a sufficient condition for
\invariantconfluence{}---dubbed \defword{\invariantclosure{}}---that is simpler to
reason about~\cite{li2012making, li2014automating}. In this section, we define
\invariantclosure{} and study why the condition is sufficient but not necessary.
Armed with this understanding, we present conditions under which it is both
sufficient and necessary.

We say an object $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to
an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying
states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
$I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.

@@

\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
sufficient for \invariantconfluence{}. Intuitively, our system model ensures that
transaction execution preserves the invariant, so if merging states also
preserves the invariant and if our start state satisfies the invariant, then
inductively it is impossible for us to reach a state that doesn't satisfy the
invariant.

This is good news because checking if an object is \invariantclosed{} is more
straightforward than checking if it is \invariantconfluent{}. Existing systems
typically use an SMT solver like Z3 to check if an object is
\invariantclosed{}~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
is, then by \thmref{IclosureImpliesIconfluence}, it is \invariantconfluent{}.
Unfortunately, \invariantclosure{} is \emph{not} necessary for
\invariantconfluence{}, so if an object is \emph{not} \invariantclosed{}, these
systems cannot conclude that the object is \emph{not} \invariantconfluent{}. The
reason why \invariantclosure{} is not necessary for \invariantconfluence{} is best
explained through an example.

@@

  The invariant and the set of reachable states are illustrated in \figref{Z2}
  in which we draw each state $(x, y)$ as a point in space. The invariant
  consists of the second and fourth quadrant, and the reachable states consist
  of the fourth quadrant. From this, it's immediate that the reachable states
  are a subset of the invariant, so $O$ is \invariantconfluent{}. However,
  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
  \invariantclosed{}. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
  (1, 1)$, we see $\lnot I(s_3)$.
\end{example}

@@

In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
\invariantclosed{}, but $s_1$ is not reachable. This is not particular to
\exampleref{Z2}. In fact, it is fundamentally the reason why \invariantclosure{}
is not equivalent to \invariantconfluence{}. \Invariantconfluence{} is, at its
core, a property of reachable states, but \invariantclosure{} is completely
ignorant of reachability. As a result, invariant satisfying yet unreachable
states like $s_1$ are the key hurdle preventing \invariantclosure{} from being
equivalent to \invariantconfluence{}. This is formalized by
\thmref{IclosureEquivalentIconfluence}

\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$. If the invariant is a subset of the
  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$), then
  \[
    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
  \]
\end{theorem}
\begin{proof}
  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
  \join s_2$, satisfy the invariant.
\end{proof}

% - define \invariantclosure{} and introduce it as a sufficient condition
% - cite existing work to show that people have used invariant closure to
%   reason about \invariantconfluence{}
% - explain why invariant closure can fail and why its not sufficient using the
%   two-integers example
% - state main theorem

### 2018-04-22 doc/paper/sections/invariant_confluence.tex
Documented segmented invariant-confluence DP.
Finished first cut at related work.
Formatting tweaks.

\section{Invariant-Confluence}\seclabel{InvariantConfluence}
Informally, a replicated object is \defword{invariant-confluent} with respect
to an invariant if every replica of the object is guaranteed to satisfy the
invariant despite the possibility of different replicas being concurrently
modified or merged together~\cite{bailis2014coordination}. In this section, we
make this informal notion of invariant-confluence precise. We begin by
introducing our system model of replicated objects and then present a formal
definition of invariant-confluence.

@@

Informally, $O$ is \defword{invariant-confluent} with respect to $s_0$, $T$, and
$I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots, s_n$
is guaranteed to always satisfy the invariant $I$ in every possible execution
of the system.

\subsection{Expression-Based Formalism}
To define invariant-confluence formally, we represent a state produced by a
system execution as a simple expression generated by the grammar
\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]
where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
an example, consider the system execution in \figref{SystemExecution} in which
a distributed object is replicated across servers $p_1$, $p_2$, and $p_3$.
Server $p_3$ begins with state $s_0$, transitions to state $s_2$ by executing
transaction $u$, transitions to state $s_5$ by executing transaction $w$, and
then transitions to state $s_7$ by merging with server $p_1$. Similarly, server
$p_1$ ends up with state $s_6$ after executing transactions $t$ and $v$ and
merging with server $p_2$. In \figref{Expression}, we see the abstract syntax
tree of the corresponding expression for state $s_7$.

@@

We say an expression $e$ is \defword{$\sTIreachable{}$} if it corresponds to a
valid execution of our system model. Formally, we define $\sTIreachable{e}$ to
be the smallest predicate that satisfies the following equations:
\begin{itemize}
  \item
    $\sTIreachable{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachable{e}$ and $I(t(e))$, then
    $\sTIreachable{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachable{e_1}$ and
    $\sTIreachable{e_2}$, then $\sTIreachable{e_1 \join e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is $\sTIreachable{}$ if there exists a
$\sTIreachable{}$ expression $e$ that evaluates to $s$.

Finally, we say $O$ is \defword{invariant-confluent} with respect to $s_0$,
$T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
satisfy the invariant: $\setst{s \in S}{\sTIreachable{s}} \subseteq I$.
% TODO: Make a note that this definition is different than the original, and
% reference the appendix where we include the proof that the two definitions
% are equivalent.

>>>>

\section{Invariant Confluence}\seclabel{InvariantConfluence}
Informally, a replicated object is \defword{\invariantconfluent{}} with respect
to an invariant if every replica of the object is guaranteed to satisfy the
invariant despite the possibility of different replicas being concurrently
modified or merged together~\cite{bailis2014coordination}. In this section, we
make this informal notion of \invariantconfluence{} precise. We begin by
introducing our system model of replicated objects and then present a formal
definition of \invariantconfluence{}.

@@

Informally, $O$ is \defword{\invariantconfluent{}} with respect to $s_0$, $T$, and
$I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots, s_n$
is guaranteed to always satisfy the invariant $I$ in every possible execution
of the system.

\subsection{Expression-Based Formalism}
To define \invariantconfluence{} formally, we represent a state produced by a
system execution as a simple expression generated by the grammar
\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]
where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
an example, consider the system execution in \figref{SystemExecution} in which
a distributed object is replicated across servers $p_1$, $p_2$, and $p_3$.
Server $p_3$ begins with state $s_0$, transitions to state $s_2$ by executing
transaction $u$, transitions to state $s_5$ by executing transaction $w$, and
then transitions to state $s_7$ by merging with server $p_1$. Similarly, server
$p_1$ ends up with state $s_6$ after executing transactions $t$ and $v$ and
merging with server $p_2$. In \figref{Expression}, we see the abstract syntax
tree of the corresponding expression for state $s_7$.

@@

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
following equations:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e$ and for all transactions $t$ in the set $T$ of
    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
    $\sTIreachablepredicate{t(e)}$.
  \item
    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
    e_2}$.
\end{itemize}
Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists a
\sTIreachable{} expression $e$ that evaluates to $s$.

Finally, we say $O$ is \defword{\invariantconfluent} with respect to $s_0$,
$T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
satisfy the invariant: $\setst{s \in S}{\sTIreachablepredicate{s}} \subseteq
I$.
